\subsection{\rqidentifying: mutation-effective metamorphic relations}

The definition and explanation of the metamorphic relations are in
Section~\ref{sec:def_MT}. In this section, we examine how each
metamorphic rule affects each part of the \lstinline|rcpsp.mzn| listed
in~\ref{lst:minizinc-rcpsp}. The constraints of the programs are
divided into four groups:

\begin{enumerate}
\item\label{item:precdence} Precedence constraints.
\item\label{item:redundant} Redundant non-overlapping constraints.
\item\label{item:cumulative} Cumulative resource constraints.
\item\label{item:makespan} Makespan constraints.
\end{enumerate}

Metamorphic rule \mr{1} detects defects in all groups except
cumulative resource constraints; \mr{2} detects defects in precedence
constraints; and \mr{3} and \mr{4} detect defects in all groups.

In terms of mutation score, the most effective rule is \mr{3},
which detects 63 mutants out of 74 mutants,
followed by \mr{4}, which detects 48
mutants, \mr{1}, which detects 24 mutants, and \mr{2}, which detects
six mutants. All the mutants detected by \mr{2} are also detected by
the other relations. Additionally, there are two mutants detected by both \mr{1}
and \mr{4} that are not detected by \mr{3}.

\mr{3} generated test cases that killed the 6 mutants that survived the initial test suite. The other metamorphic relations did not produce tests that killed those surviving mutants. % interpretación: estudiar por qué MR3 específicamente mata a esos seis mutantes - qué hay de especial en las pruebas que genera, y cómo se relacionan esas pruebas con los seis mutantes supervivientes?

% Esta tabla no tiene sentido. Los 6 mutantes supervivientes
% solo son detectados por MR3
% \begin{table}[]
%     \centering
%     \begin{tabular}{l r rrrrrr}
%     \toprule
%          MR & \# killed & $m_1$ & $m_2$ & $m_3$ & $m_4$ & $m_5$ & $m_6$ \\
%          \midrule
%          % TODO: rellenar tabla
%          \mr{1} & 24 & 1 & 1 & 0 & 0 & 0 & 0 \\
%          \mr{2} & 24 & 1 & 1 & 0 & 0 & 0 & 0 \\
%          \mr{3} & 24 & 1 & 1 & 0 & 0 & 0 & 0 \\
%          \mr{4} & 24 & 1 & 1 & 0 & 0 & 0 & 0 \\
%          \bottomrule
%     \end{tabular}
%     \caption{Number of mutants killed by the follow-up test cases from each metamorphic rule, and whether they killed (1) or not (0) the 6 mutants that survived the initial test suite ($m_1 \cdots m_6$.}
%     \label{tab:mr-comparison}
% \end{table}


%%% Local Variables:
%%% mode: LaTeX
%%% TeX-master: "MT_scheduling"
%%% End:
