
\section{Conclusions and  future work}\label{sec6}

%%%Conclusions

We have proposed a metamorphic testing framework for scheduling programs in the 
MiniZinc constraint language. The metamorphic relations have been designed, 
the corresponding follow-up test cases have been generated. 
In order to validate the MRs, the original code has been mutated by 
applying several mutation operators. They have been validated with promising results. 

Furthermore, these experiments have allowed, on the one hand, 
to design new mutation operators specific to declarative languages. 
On the other hand, a bug has been detected in a version of the solver. 
These discoveries demonstrate the usefulness of metamorphic testing,
because of the new perspective it brings compared to other software
testing techniques.

%Hemos testeado por primera vez un lenguaje y sistema que es puntero en la programaci'on con restricciones
%Hemos comprobado que nuestros test son v'alidos pues hemos detectado un error en una versi'on antigua que en la ersi'on actual ya est'a arreglado.
%Hemos utilizado herramientas nuevas como testear si los par'ametros de una funci'on est'an mal colocados. Hay que tener en cuenta que las funciones globales hay mucho par'ametros que son listas (arrays) de enteros.


% Future work

Future work includes further automation of the process, design of new MRs
and, application to other types of problems such as supply chain management, and transportation.
%Automatizar todo esto
%Buscar m'as MTs
%LLevar estas ideas a otros tipos de problemas como supply chain man- agement, and transportation,
In addition, the set of mutation operators for
this language is not fully developed, which is a line
to be deployed in the future.

