\section{Related work}\label{sec2}

 %\subsection{Constraint solving and MT}
%%%  2003: Automated Metamorphic Testing       DBLP:conf/compsac/GotliebB03
%%% MT+CP "Metamorphic testing of constraint solvers"   akgun2018metamorphic

Constraint solving and metamorphic testing have been combined in different ways in the past. However, existing work has focused on other aspects or domains and their approaches differ from ours. First, Gotlieb and Botella in~\cite{DBLP:conf/compsac/GotliebB03} use constraint logic programming to find test cases that fail a given MR. They present some examples where it can be demonstrated that the program fulfills these MRs. The paper presents a prototype implementation of their approach using the existing INKA test case generator~\cite{ingenierie2002thales}.

On the other hand, a recently published paper ~\cite{akgun2018metamorphic} proposes to use MT to validate constraint programming systems. The complexity of these systems makes them difficult to test otherwise. In particular, they evaluate the use of MT to validate the Minion CPS~\cite{gent2006minion}. They show that MT is effective in finding artificial errors introduced by random code mutations.

%\subsection{Scheduling and MT}
%%% MT+Scheduling
%%% 2013: Testing Central Processing Unit Scheduling Algorithms Using Metamorphic Testing
%%% Mingyue Jiang et al.
%%% Jiang13MT
Regarding scheduling problems in relation to MT, Jiang et al. in ~\cite{Jiang13MT}, apply MT to an algorithm for Central Processing Unit (CPU) scheduling programs. Although they also deal with testing algorithms for managing processes (with resource constraints), unlike our proposal, they limit themselves to testing an algorithm used for managing CPU processes. In our case, the context is totally different, as well as the code to be tested.

%%%% Añadir esta referencia:
%%% XIE2013866
%%% 2013: Metamorphic slice: An application in spectrum-based fault localization
%%% uTILIZAN LA SUITE DE SIEMENS QUE TIENE UN PAR DE PROGRAMAS DENOMINADOS SCHEDULING Y SCHEDULIN2
%%% El objetivo es diferente del nuestro pero usan los progrmas para probar programas 
%%% sin oráculos con MT
%%% Añadir texto




Furthermore, constraint programming is used in solving NP-hard scheduling problems.
The success of constraint-based programming in real projects is due to the combination of the techniques of two research areas: Operations Research (OR) and Artificial Intelligence (AI) \cite{bartak2010constraint}.
%
Scheduling is applied in a variety of areas, such as assigning gates to planes arriving at an airport, workers to an assembly line, or processes to a CPU or multiple CPUs.
On the other hand, the volume of research on scheduling problems indicates how important this area is in solving real-life problems.
We refer to the surveys provided by
Icmeli et al. \cite{icmeli1993project},
Özdamar and Ulusoy \cite{ozdamar1995survey},
Herroelen et al. \cite{herroelen1998resource},
Brucker et al. \cite{brucker1999resource},
Weglarz \cite{weglarz2012project} and, most recent, 
%Hartmann and Kolisch \cite{hartmann2000experimental} and, 
Hartmann et al. \cite{hartmann2022updated}.
